### 2023-03-15 article3.qmd
update

@@

We applied a similar approach to @Estapa2013-gw and @Estapa2017-jr to derive the continuous component (i.e. continuous deposition of small particles on the upward-facing window of the transmissometer) and discontinuous (i.e. deposition of stochastic big particles) components of the downward POC flux, hereafter referred to as, respectively, $F_{small}$ and $F_{big}$.

@@

Contrarily to previous works [@Estapa2013-gw; @Estapa2017-jr; @Estapa2023-oz], we chose an empirical statistical approach instead of fixed thresholds to detect jumps. Therefore, we firstly computed the slope between each $c_{p}$ observation for each drifting phase and computed the z-score (i.e. standard score) of each of them. The detection of jumps therefore consisted in spotting outliers in the z-score distribution. For this, we computed the interquartile range ($IQR$) between the first ($Q_{1}$) and third ($Q_{3}$) quartiles. Positive and negative jumps were defined as, respectively, a z-score above $Q_{3} + 1.5 \times IQR$ and below $Q_{1} - 1.5 \times IQR$. We found that this method better detected outliers than the one used to despike the $c_{p}$ signal.

@@

We used the same relationship to compute $F_{big}$ by replacing $slope_{w}$ as the sum of positive jumps divided by the drifting time.

@@

# add big particles flux
rows_to_keep <- c(jump_index, jump_index-1) 
tmp3 <- tmp2[rows_to_keep,] |> select(dates, cp, jump, colour, group) |> dplyr::arrange(dates)

@@

# big particles flux
if(is.null(tmp3)){ # no jumps
  big_part_poc_flux <- 0
}else{
  tmp4 <- tmp3 |> filter(diff_jump > 0)
  if(nrow(tmp4) == 0){ # no positive jumps
    big_part_poc_flux <- 0
  }else{
    delta_y <- sum(tmp4$diff_jump) *0.25 # to get ATN (= cp*0.25)
    max_time <- max(tmp2$dates)
    min_time <- min(tmp2$dates)
    delta_x <- as.numeric(difftime(max_time, min_time, units = 'days'))
    slope_big_part <- delta_y/delta_x
    big_part_poc_flux <- 633*(slope_big_part**0.77)
  }
}

@@

@@

  # add big particles flux to the party
  rows_to_keep <- c(jump_index, jump_index-1)
  tmp2 <- tmp[rows_to_keep,] |> select(dates, cp, slope, colour, group) |> arrange(dates)

@@

  if(is.null(tmp2)){ # no jump
    big_part_poc_flux <- 0
    tmp3 <- NULL
  }else{
    tmp3 <- tmp2 |> filter(diff_jump > 0)
    if(nrow(tmp3) == 0){ # no positive jumps
      big_part_poc_flux <- 0
    }else{
      delta_y <- sum(tmp3$diff_jump) *0.25 # to get ATN (= cp*0.25)
      max_time <- max(tmp$dates)
      min_time <- min(tmp$dates)
      delta_x <- as.numeric(difftime(max_time, min_time, units = 'days'))
      slope_big_part <- delta_y/delta_x
      big_part_poc_flux <- 633*(slope_big_part**0.77)
    }
  }

@@

  # plot
  jump_plot <- plot_ly(tmp, x = ~dates, y = ~cp, type = 'scatter', mode = 'markers', color = ~colour, colors = c('#003366','#E31B23', '#FFC325')) |>
    add_lines(data= part_slope_tmp, x = ~dates, y = ~cp, split = ~group, color = I('#DCEEF3'), showlegend = F) |>
    layout(title= paste0('Drifting time: ', round(drifting_time,3), ' days\n',
                         'Mean ATN slope (light blue): ', round(mean_slope,3), ' day-1\n',
                         'POC flux (small particles): ', round(poc_flux,1), ' mg C m-2 day-1\n',
                         'POC flux (big particles): ', round(big_part_poc_flux,1), ' mg C m-2 day-1'), yaxis = list(title = 'Cp (1/m)'), xaxis = list(title = 'Time'))

@@

    # adapt script to return big and small flux
  df <- tibble('max_time' = max_time, 'min_time' = min_time, 'small_flux' = poc_flux, 'big_flux' = big_part_poc_flux, park_depth = data$park_depth[1], wmo = data$wmo[1],
               cycle = data$cycle[1])

@@

```{r, compute Fsmall and F big}

@@

cflux_info_table <- cflux |> dplyr::group_by(park_depth) |> summarize(min_smallf = min(small_flux, na.rm=T), max_smallf = max(small_flux, na.rm=T),
                                                                mean_smallf = mean(small_flux, na.rm=T), median_smallf = median(small_flux, na.rm=T),
                                                                std_smallf = sd(small_flux, na.rm=T),
                                                                min_bigf = min(big_flux, na.rm=T), max_bigf = max(big_flux, na.rm=T),
                                                                mean_bigf = mean(big_flux, na.rm=T), median_bigf = median(big_flux, na.rm=T),
                                                                std_bigf = sd(big_flux, na.rm=T))

@@

$F_{small}$ and $F_{big}$ were computed at each drifting depth for all floats (@fig-OST-flux). At 200 m, $F_{small}$ ranges between `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$min_smallf, 1)` and `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$max_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$mean_smallf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$std_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ (average $\pm$ standard deviation). At 500 m, $F_{small}$ is comprised between `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$min_smallf, 1)` and `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$max_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$mean_smallf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$std_smallf, 1)` mg C m$^{2}$ day$^{-1}$. For both drifting depths, the trend shows a decrease of $F_{small}$ from June to, respectively, January and February before the drift was stopped at those depths due to the deepening of the winter MLD (@fig-mld) in order to ensure the float stability. At 1000 m, all floats measured an increase of $F_{small}$ from late May to mid/end of July followed by a steady decrease towards March, with $F_{small}$ ranging between `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$min_smallf, 1)` and `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$max_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ and `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$mean_smallf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$std_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ on average.

```{r, Fsmall and Fbig, fig.width=8, fig.height=6}
#| label: fig-OST-flux
#| fig-cap: OST-derived small (top) and big (bottom) particles carbon flux for all floats at each drifting depth. Black lines and shaded areas represent, respectively, the monthly mean flux and the 95% confidence interval around the mean. There is no data after January and February at, respectively, 200 m and 500 m because the drifting was stopped at those depths due to the deepening of the MLD.

# add black data to have the same x axis for all drifting depths (to show that we stopped the drifting at 200 and 500 m)
blank_data <- tibble(max_time = c(NA, NA), min_time = c(NA,NA), small_flux = c(NA,NA), big_flux = c(NA,NA), park_depth = c('200 m', '500 m'),
                     wmo = rep(4903634,2), cycle = c(NA,NA), drifting_time = c(NA,NA), WMO = rep(as.factor(4903634),2), date = rep('2023-03-05',2))

@@

down <- tmp |> ggplot(aes(x = date, y = big_flux)) +
  geom_point(data = tmp, aes(x = date, y = big_flux, colour = WMO, shape = WMO)) +
  geom_smooth(data = tmp, aes(x = date, y = big_flux), colour = 'black') +
  scale_color_brewer(palette = 'Dark2') + 
  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
  guides(fill=guide_legend(title="WMO")) +
  facet_wrap(~park_depth, scales = 'free') +
  labs(x = 'Month', y = TeX('$F_{big}$ (mg C m$^{-2}$ day$^{-1}$)')) +
  theme_bw() +
  scale_y_continuous(trans = 'log10') +
  theme(legend.position = 'none')

@@

$F_{big}$ was estimated at 200 m between 0 (no deposition of big particles on the transmissometer window) and `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$max_bigf, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$mean_bigf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$std_bigf, 1)` mg C m$^{-2}$ day$^{-1}$. High fluxes (\> 100 mg C m$^{-2}$ day$^{-1}$) were measured by 3 out of 4 floats at the end of May before a sharp decrease towards July. At 500 m, $F_{big}$ ranges between `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$min_bigf, 1)` mg C m$^{-2}$ day$^{-1}$ and `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$max_bigf, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$mean_bigf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$std_bigf, 1)` mg C m$^{-2}$ day$^{-1}$, decreasing from September to February. At 1000 m, $F_{big}$ shows a similar pattern than the one observed for $F_{small}$ with an increase from June to mid-July, followed by a decrease towards March, with values ranging from 0 to `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$max_bigf, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$mean_bigf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$std_bigf, 1)` mg C m$^{-2}$ day$^{-1}$.

@@

Vertical profiles of particle concentration in the lower (102 - 512 $\mu$m) and upper (0.512 - 2.5 mm) size range of the UVP6 shows similar patterns across all floats. In the lower part of the size spectrum, @fig-uvp6-profiling-102-406 shows the rapid emptying of the water column in November. This process occurred on the entire water column (e.g. size class 406, float 4903634) but also started with more intensity from below the surface (e.g. size class 102, float 6904240). All floats also measured a local particle abundance maximum around 800 - 1000 m whose origin could be due to a previous particle export or remineralization of bigger particles.

@@

```{r, UVP6 profiling mode without sinking speed big classes, fig.height=8, fig.width=8}
#| label: fig-uvp6-profiling-512-2500
#| fig-cap: Vertical profiles of UVP6 particle concentration for size classes in the 0.512 - 2.5 mm size range. Concentrations were normalized for each size class. The colorbar is in log$_{10}$ scale. The gray background depicts the absence of particle in the specified size class.

@@

The Chla satellite pattern for the 6904240-4903634 pair typically describes the dynamics of the spring bloom in the northern part of the LS [@Lacour2015-da]. This is not surprising given that this pair was deployed one degree southward to the so-called 60°N limit [@Lacour2015-da]. @fig-chla-mld-satellite shows that the satellite MLD in that region had already shoaled below 100 m since March and that the water column was stable and relatively well stratified in May which might explain the intensity of surface Chla concentrations in contrast with those measured at the second deployment site. The two other floats were deployed in the bioregion where the phytoplankton biomass typically develops slowly in May to reach its maximum in June [@Lacour2015-da]. However, the Chla pattern observed in @fig-chla-mld-satellite could suggest that the development of the phytoplankton biomass was stopped by an increase in the MLD at the end of May and that it started again in mid-June when the water column was well stratified. <!-- However, the increase of the MLD in June could have also impacted the intensity of the bloom which might explain the three times lower than Chla concentrations. However, the Chla satellite concentration (before deployment) for that pair was three times lower than for the other pair and the second Chla increase (satellite and in situ) was of the same order of magnitude. Therefore, the second pair was deployed in a less productive area and that the inten <!--Hence, it is possible that the bloom intensity was not sufficient to produce clear particle export signals, at least for BGC-Argo floats in their 10-day vertical sampling configuration. -->

@@

Here, we refer to the first and second deployment as, respectively, station A and station B. High Performance Liquid Chromatography (HPLC) analyses were performed on samples taken during the two deployments. The selected pigments, their taxonomic significance and their size class were taken from Table 2 in @Petit2022-de. Station A showed Chla concentrations (TChla pigment in @fig-hplc) two times higher than in station B in the first 100 m as well as higher concentrations of diatoms (Fucoxanthin). Station B showed higher concentrations of cryptophytes (Alloxanthin), dinoflagellates (Peridin) and green flagellates and prochlorophytes (TChlb) than in station A. However, their absolute concentrations are quite low (usually less than 0.05 mg m$^{-3}$). Taking that into consideration, it seems that station A was more productive than station B and was dominated by diatoms (i.e. microphytoplankton) whereas station B was composed of a majority of smaller phytoplankton species (mostly nano and picozooplankton). This supports the idea of two distinct bloom dynamics as inferred from satellite data.

@@

ggplot(hplc) + geom_path(aes(y = depth, x = concentration, colour = station)) + 
  facet_wrap(~pigment, scales = 'free') + scale_y_reverse() +
  labs(colour = 'Station', x = TeX('Concentration (mg m$^{-3}$)'), y = 'Depth (m)') +
  theme_bw() + scale_color_manual(values = c('#E31B23', '#003366'))
```

@@

Even though we suggested that the two pairs of floats were deployed in two different bloom regimes, they all measured an increase in particle concentrations in the 102 - 512 $\mu$m range (@fig-uvp6-profiling-102-406) around 800 - 1000 m as well as an increase in the abundance of big particles (@fig-uvp6-profiling-512-2500) between June and August. This increase raises the question of a possible particle export event. For classes \> 512 $\mu$m, the vertical concentration profiles highlight a bigger particle plume for the 6904241-1902578 pair than for the other one. However, apart from the float 6904240, all floats measured similar integrated Chla content in the upper 100 m and in the mixed layer (@fig-chla-content). @fig-OST-flux also shows that $F_{small}$ is similar across all floats at all parking depths and that $F_{big}$ is also similar across all floats at 1000 m with the same increase (from June to August) than for F$_{small}$. The same behavior is observed for UVP6 particles in the lower size range at 1000 m (@fig-uvp-1000m). Observing similar properties across a wide geographical range around deployments suggests that we observe the latent dynamics of the system rather than episodic export events.

@@

# test taxo -> pas forcément concluant -> à creuser..
#ggplot(taxo_1902578 |> filter(cycle == 3)) + geom_point(aes(x = concentration, y = -depth))
```

@@

taxo_plot <- all_taxo |> filter(depth <= 1000) |> mutate(date = as.Date(date)) |> filter(date < as.Date("2023-06-03")) |>
  #group_by(size) |>
  ggplot() + facet_wrap(~wmo, ncol = 4) + theme_bw() +
  geom_point(aes(x=date, y=depth, colour=log10(concentration)), size=1, shape=15) +
  scale_color_cmocean(name = 'thermal', limits = c(-2.5,-1), oob = scales::squish, na.value = 'grey', direction = 1) +
  scale_y_reverse() +
  scale_x_date(labels = date_format("%m"), date_breaks = '3 month') +
  labs(colour = 'Calanoida (#/L)    ', x = 'Month', y = 'Depth (m)') + theme(axis.title.x = element_blank())

@@

For the first time, BGC-Argo floats equipped with both a transmissometer and a UVP6 have been deployed to better characterize the downward carbon flux and the nature of sinking particles from a multi-instrument approach [@Claustre2021-tx]. The quantification of the POC flux using the OST method [@Bishop2004-at; @Estapa2013-gw; @Estapa2017-jr] is based on an empirical POC:ATN ratio \[REF estapa 2022\] whereas UVP6 PSD measurements are used to compute POC fluxes using @eq-guidi-article3:

@@

where $C_{i}$, A and b represent, respectively, the concentration of particles (number/L) for the $i^{th}$ class of mean diameter $d_{i}$ (in mm) and 2 constants with $b$ being related to the fractal dimension ($D$) of an aggregate by $D = (b +1)/2$ [@Guidi2008-vp].

@@

```{r, compute guidi fluxes at parking depth, fig.height=6, fig.width=6}
#| label: fig-comparison-ost-uvp
#| fig-cap: Comparison of total carbon fluxes derived from the OST (black) and the UVP6 (red) where the error bars represent the 25$^{th}$ and 75$^{th}$ percentiles around the median.

>>>>

In the first chapter, I explained the different steps that led to the creation of the embedded zooplankton classification model for the UVP6. However, I did not specify that in spite of the technical and energetic constraints, the model was also due for April 2022 for a deployment of BGC-Argo floats in the Labrador Sea a month later. Once deployed, it was only a question of time before I could dig into the UVP6 and the OST data. In the meantime I asked myself, what could be a cool thing to do about those data? I already had some experience with BGC-Argo floats during my master's degree (study of the deep chlorophyll maximum in the Black Sea, see also the abstract of that paper in APPENDIX XX) and because I was a bit bored waiting for new BGC-Argo profiles, I thought it would be useful to have a visualization tool in the form of a web app to investigate incoming data and already prepare the ground for the following study. Here is the link to the [web app](https://fricour.shinyapps.io/CARBON_REFINE/), which is still currently being updated every night with new data (near) live from the Labrador Sea and hopefully with other BGC-Argo floats equipped with the UVP6 and the OST to investigate carbon and particle fluxes in other regions. This study uses preliminary data (not even one year of data) and a new BGC-Argo float configuration hence more work should be addressed to better understand the dynamics of the system.

@@

We applied a similar approach to @Estapa2013-gw and @Estapa2017-jr to derive the continuous component (i.e. continuous deposition of small particles on the upward-facing window of the transmissometer) and discontinuous (i.e. deposition of stochastic large particles) components of the downward POC flux, hereafter referred to as, respectively, $F_{small}$ and $F_{large}$.

@@

Contrarily to previous works [@Estapa2013-gw; @Estapa2017-jr; @Estapa2023-oz], we chose an empirical statistical approach instead of fixed thresholds to detect jumps. Therefore, we firstly computed the slope between each $c_{p}$ observation for each drifting phase and computed the z-score (i.e. standard score) of each of them. The detection of jumps therefore consisted in spotting outliers in the slope distribution. For this, we computed the interquartile range ($IQR$) between the first ($Q_{1}$) and third ($Q_{3}$) quartiles. Positive and negative jumps were defined as, respectively, a slope above $Q_{3} + 1.5 \times IQR$ and below $Q_{1} - 1.5 \times IQR$. We found that this method better detected outliers than the one used to despike the $c_{p}$ signal.

@@

We used the same relationship to compute $F_{large}$ by replacing $slope_{w}$ as the sum of positive jumps divided by the drifting time.

@@

# add large particles flux
rows_to_keep <- c(jump_index, jump_index-1) 
tmp3 <- tmp2[rows_to_keep,] |> select(dates, cp, jump, colour, group) |> dplyr::arrange(dates)

@@

# large particles flux
if(is.null(tmp3)){ # no jumps
  large_part_poc_flux <- 0
}else{
  tmp4 <- tmp3 |> filter(diff_jump > 0)
  if(nrow(tmp4) == 0){ # no positive jumps
    large_part_poc_flux <- 0
  }else{
    delta_y <- sum(tmp4$diff_jump) *0.25 # to get ATN (= cp*0.25)
    max_time <- max(tmp2$dates)
    min_time <- min(tmp2$dates)
    delta_x <- as.numeric(difftime(max_time, min_time, units = 'days'))
    slope_large_part <- delta_y/delta_x
    large_part_poc_flux <- 633*(slope_large_part**0.77)
  }
}

@@

  # IQR <- quantile(tmp$slope, probs = 0.75, na.rm=T) - quantile(tmp$slope, probs = 0.25, na.rm=T)
  # # outliers ('spikes' in the Z score signal)
  # spikes_down <- tmp$slope < quantile(tmp$slope, 0.25, na.rm=T) - 1.5 *IQR
  # spikes_up <-  tmp$slope > quantile(tmp$slope, 0.75, na.rm=T) + 1.5 *IQR
  # spikes <- as.logical(spikes_down + spikes_up)

@@

  # add large particles flux to the party
  rows_to_keep <- c(jump_index, jump_index-1)
  tmp2 <- tmp[rows_to_keep,] |> select(dates, cp, slope, colour, group) |> arrange(dates)

@@

  if(is.null(tmp2)){ # no jump
    large_part_poc_flux <- 0
    tmp3 <- NULL
  }else{
    tmp3 <- tmp2 |> filter(diff_jump > 0)
    if(nrow(tmp3) == 0){ # no positive jumps
      large_part_poc_flux <- 0
    }else{
      delta_y <- sum(tmp3$diff_jump) *0.25 # to get ATN (= cp*0.25)
      max_time <- max(tmp$dates)
      min_time <- min(tmp$dates)
      delta_x <- as.numeric(difftime(max_time, min_time, units = 'days'))
      slope_large_part <- delta_y/delta_x
      large_part_poc_flux <- 633*(slope_large_part**0.77)
    }
  }

@@

  # plot
  jump_plot <- plot_ly(tmp, x = ~dates, y = ~cp, type = 'scatter', mode = 'markers', color = ~colour, colors = c('#003366','#E31B23', '#FFC325')) |>
    add_lines(data= part_slope_tmp, x = ~dates, y = ~cp, split = ~group, color = I('#DCEEF3'), showlegend = F) |>
    layout(title= paste0('Drifting time: ', round(drifting_time,3), ' days\n',
                         'Mean ATN slope (light blue): ', round(mean_slope,3), ' day-1\n',
                         'POC flux (small particles): ', round(poc_flux,1), ' mg C m-2 day-1\n',
                         'POC flux (large particles): ', round(large_part_poc_flux,1), ' mg C m-2 day-1'), yaxis = list(title = 'Cp (1/m)'), xaxis = list(title = 'Time'))

@@

    # adapt script to return large and small flux
  df <- tibble('max_time' = max_time, 'min_time' = min_time, 'small_flux' = poc_flux, 'large_flux' = large_part_poc_flux, park_depth = data$park_depth[1], wmo = data$wmo[1],
               cycle = data$cycle[1])

@@

```{r, compute Fsmall and F large}

@@

cflux_info_table <- cflux |> dplyr::group_by(park_depth) |> summarize(min_smallf = min(small_flux, na.rm=T), max_smallf = max(small_flux, na.rm=T),
                                                                mean_smallf = mean(small_flux, na.rm=T), median_smallf = median(small_flux, na.rm=T),
                                                                std_smallf = sd(small_flux, na.rm=T),
                                                                min_largef = min(large_flux, na.rm=T), max_largef = max(large_flux, na.rm=T),
                                                                mean_largef = mean(large_flux, na.rm=T), median_largef = median(large_flux, na.rm=T),
                                                                std_largef = sd(large_flux, na.rm=T))

@@

$F_{small}$ and $F_{large}$ were computed at each drifting depth for all floats (@fig-OST-flux). At 200 m, $F_{small}$ ranges between `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$min_smallf, 1)` and `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$max_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$mean_smallf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$std_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ (average $\pm$ standard deviation). At 500 m, $F_{small}$ is comprised between `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$min_smallf, 1)` and `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$max_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$mean_smallf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$std_smallf, 1)` mg C m$^{2}$ day$^{-1}$. For both drifting depths, the trend shows a decrease of $F_{small}$ from June to, respectively, January and February before the drift was stopped at those depths due to the deepening of the winter MLD (@fig-mld) in order to ensure the float stability. At 1000 m, all floats measured an increase of $F_{small}$ from late May to mid/end of July followed by a steady decrease towards March, with $F_{small}$ ranging between `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$min_smallf, 1)` and `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$max_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ and `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$mean_smallf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$std_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ on average.

```{r, Fsmall and Flarge, fig.width=8, fig.height=6}
#| label: fig-OST-flux
#| fig-cap: OST-derived small (top) and large (bottom) particles carbon flux for all floats at each drifting depth. Black lines and shaded areas represent, respectively, the monthly mean flux and the 95% confidence interval around the mean. There is no data after January and February at, respectively, 200 m and 500 m because the drifting was stopped at those depths due to the deepening of the MLD.

# add black data to have the same x axis for all drifting depths (to show that we stopped the drifting at 200 and 500 m)
blank_data <- tibble(max_time = c(NA, NA), min_time = c(NA,NA), small_flux = c(NA,NA), large_flux = c(NA,NA), park_depth = c('200 m', '500 m'),
                     wmo = rep(4903634,2), cycle = c(NA,NA), drifting_time = c(NA,NA), WMO = rep(as.factor(4903634),2), date = rep('2023-03-05',2))

@@

down <- tmp |> ggplot(aes(x = date, y = large_flux)) +
  geom_point(data = tmp, aes(x = date, y = large_flux, colour = WMO, shape = WMO)) +
  geom_smooth(data = tmp, aes(x = date, y = large_flux), colour = 'black') +
  scale_color_brewer(palette = 'Dark2') + 
  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
  guides(fill=guide_legend(title="WMO")) +
  facet_wrap(~park_depth, scales = 'free') +
  labs(x = 'Month', y = TeX('$F_{large}$ (mg C m$^{-2}$ day$^{-1}$)')) +
  theme_bw() +
  scale_y_continuous(trans = 'log10') +
  theme(legend.position = 'none')

@@

$F_{large}$ was estimated at 200 m between 0 (no deposition of large particles on the transmissometer window) and `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$max_largef, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$mean_largef, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$std_largef, 1)` mg C m$^{-2}$ day$^{-1}$. High fluxes (\> 100 mg C m$^{-2}$ day$^{-1}$) were measured by 3 out of 4 floats at the end of May before a sharp decrease towards July. At 500 m, $F_{large}$ ranges between `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$min_largef, 1)` mg C m$^{-2}$ day$^{-1}$ and `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$max_largef, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$mean_largef, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$std_largef, 1)` mg C m$^{-2}$ day$^{-1}$, decreasing from September to February. At 1000 m, $F_{large}$ shows a similar pattern than the one observed for $F_{small}$ with an increase from June to mid-July, followed by a decrease towards March, with values ranging from 0 to `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$max_largef, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$mean_largef, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$std_largef, 1)` mg C m$^{-2}$ day$^{-1}$.

@@

Vertical profiles of particle concentration in the lower (102 - 512 $\mu$m) and upper (0.512 - 2.5 mm) size range of the UVP6 shows similar patterns across all floats. In the lower part of the size spectrum, @fig-uvp6-profiling-102-406 shows the rapid emptying of the water column in November. This process occurred on the entire water column (e.g. size class 406, float 4903634) but also started with more intensity from below the surface (e.g. size class 102, float 6904240). All floats also measured a local particle abundance maximum around 800 - 1000 m whose origin could be due to a previous particle export or remineralization of largeger particles.

@@

```{r, UVP6 profiling mode without sinking speed large classes, fig.height=8, fig.width=8}
#| label: fig-uvp6-profiling-512-2500
#| fig-cap: Vertical profiles of UVP6 particle concentration for size classes in the 0.512 - 2.5 mm size range. Concentrations were normalized for each size class. The colorbar is in log$_{10}$ scale. The gray background depicts the absence of particle in the specified size class.

@@

The Chla satellite pattern for the 6904240-4903634 pair typically describes the dynamics of the spring bloom in the northern part of the LS [@Lacour2015-da]. This is not surprising given that this pair was deployed one degree southward to the so-called 60°N limit [@Lacour2015-da]. @fig-chla-mld-satellite shows that the satellite MLD in that region had already shoaled below 100 m since March and that the water column was stable and relatively well stratified in May which might explain the intensity of surface Chla concentrations in contrast with those measured at the second deployment site. The latter is located in the bioregion where the phytoplankton biomass typically develops slowly in May to reach its maximum in June [@Lacour2015-da]. However, the Chla pattern observed in @fig-chla-mld-satellite could suggest that the development of the phytoplankton biomass was stopped by an increase in the MLD at the end of May and that it started again in mid-June when the water column was well stratified. <!-- However, the increase of the MLD in June could have also impacted the intensity of the bloom which might explain the three times lower than Chla concentrations. However, the Chla satellite concentration (before deployment) for that pair was three times lower than for the other pair and the second Chla increase (satellite and in situ) was of the same order of magnitude. Therefore, the second pair was deployed in a less productive area and that the inten <!--Hence, it is possible that the bloom intensity was not sufficient to produce clear particle export signals, at least for BGC-Argo floats in their 10-day vertical sampling configuration. -->

@@

Here, we refer to the first and second deployment sites as, respectively, station A and station B. High Performance Liquid Chromatography (HPLC) analyses were performed on samples taken during the two deployments. The selected pigments, their taxonomic significance and their size class were taken from Table 2 in @Petit2022-de. Station A showed Chla concentrations (TChla pigment in @fig-hplc) two times higher than in station B in the first 100 m as well as higher concentrations of diatoms (Fucoxanthin). Station B showed higher concentrations of cryptophytes (Alloxanthin), dinoflagellates (Peridin) and green flagellates and prochlorophytes (TChlb) than in station A. However, their absolute concentrations are quite low (usually less than 0.05 mg m$^{-3}$). Taking that into consideration, it seems that station A was more productive than station B and was dominated by diatoms (i.e. microphytoplankton) whereas station B was composed of a majority of smaller phytoplankton species (mostly nano and picozooplankton). This supports the idea of two distinct bloom dynamics as inferred from float and satellite data.

@@

ggplot(hplc) + geom_path(aes(y = depth, x = concentration, colour = station)) + 
  facet_wrap(~pigment, scales = 'free') + scale_y_reverse() +
  labs(colour = '', x = TeX('Concentration (mg m$^{-3}$)'), y = 'Depth (m)') +
  theme_bw() + scale_color_manual(values = c('#E31B23', '#003366'))
```

@@

Even though we suggested that the two pairs of floats were deployed in two different bloom regimes, they all measured an increase in particle concentrations in the 102 - 512 $\mu$m range (@fig-uvp6-profiling-102-406) around 800 - 1000 m as well as an increase in the abundance of large particles (@fig-uvp6-profiling-512-2500) between June and August. This increase raises the question of a possible particle export event. For classes \> 512 $\mu$m, the vertical concentration profiles highlight a largeger particle plume for the 6904241-1902578 pair than for the other one. However, apart from the float 6904240, all floats measured similar integrated Chla content in the upper 100 m and in the mixed layer (@fig-chla-content). @fig-OST-flux also shows that $F_{small}$ is similar across all floats at all parking depths and that $F_{large}$ is also similar across all floats at 1000 m with an increase from June to August as observed in F$_{small}$. The same behavior is observed for UVP6 particles in the lower size range at 1000 m (@fig-uvp-1000m). Observing similar properties across a wide geographical range around deployments suggests that we observe the latent dynamics of the system rather than episodic export events.

@@

# test taxo -> pas forcément concluant -> à creuser..
#ggplot(taxo_6904241 |> filter(cycle == 3)) + geom_point(aes(x = conc_calanoid, y = -depth))
```

@@

taxo_plot <- all_taxo |> filter(depth <= 1000) |> mutate(date = as.Date(date)) |> filter(date < as.Date("2023-06-03")) |>
  #group_by(size) |>
  ggplot() + facet_wrap(~wmo, ncol = 4) + theme_bw() +
  geom_point(aes(x=date, y=depth, colour=log10(conc_calanoid)), size=1, shape=15) +
  scale_color_cmocean(name = 'thermal', limits = c(-2.5,-1), oob = scales::squish, na.value = 'grey', direction = 1) +
  scale_y_reverse() +
  scale_x_date(labels = date_format("%m"), date_breaks = '3 month') +
  labs(colour = 'Calanoida (#/L)    ', x = 'Month', y = 'Depth (m)') + theme(axis.title.x = element_blank())

@@

#### Carbon flux trends and estimations

For the first time, BGC-Argo floats equipped with both an OST and a UVP6 have been deployed to better characterize the downward carbon flux and the nature of sinking particles from a multi-instrument approach [@Claustre2021-tx]. The quantification of the POC flux using the OST method [@Bishop2004-at; @Estapa2013-gw; @Estapa2017-jr; @Estapa2019-nk; @Estapa2023-oz] is based on an empirical POC:ATN ratio [@Estapa2023-oz] whereas UVP6 PSD measurements are used to compute POC fluxes using @eq-guidi-article3:

@@

where $C_{i}$, A and b represent, respectively, the concentration of particles (#/L) for the $i^{th}$ class of mean diameter $d_{i}$ (in mm) and 2 constants with $b$ being related to the fractal dimension ($D$) of an aggregate by $D = (b +1)/2$ [@Guidi2008-vp].

@@

```{r, compute guidi fluxes at parking depth, fig.height=6, fig.width=8}
#| label: fig-comparison-ost-uvp
#| fig-cap: Comparison of total carbon fluxes (F$_{small}$ + F$_{large}$) derived from the OST (black) and the UVP6 using the A and b values of @Guidi2008-vp (red) and @Iversen2010-an (yellow) in @eq-guidi-article3. OST data from float 6904241 stops in November due to a transmissometer failure.
